Developing effective theories for single and bilayer graphene using density matrix downfolding

### 2019-08-10 draft.tex
updated commitee
rearranged last paragraph of intro, based on joaos comments

\begin{document}
\maketitle
\begin{abstract}
A core problem of condensed matter physics is understanding how to develop low-energy effective models for realistic systems.
I believe that a density matrix downfolding (DMD) method \cite{Zheng2017} can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian.
I present as a preliminary result an effective Hamiltonian developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
I propose using DMD to develop accurate model Hamiltonians for single and bilayer graphene.
My work will serve to mature the DMD method and take steps towards a model for the strongly correlated twisted bilayer graphene system.
\end{abstract}
Committee: Lucas Wagner, Smitha Vishveshwara, David Ceperley, Karin Dahmen

Pool: \textit{David Ceperley},\textit{ Brian DeMarco}, Smitha Vishveshwara, Andre Schleife, \textit{Karin Dahmen} 
\pagebreak

@@

I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
Recent observations of a fractional quantum Hall effect \cite {Bolotin2009, Du2009} and plasmarons \cite{Bostwick999} in SLG as well as unconventional superconductivity \cite{Cao2018_2} and many-body insulating states \cite{Cao2018} in TBLG have spurred an interest in developing many-body models for graphene.
The DMD procedure provides a framework for developing such a many-body model using QMC calculations starting from the \textit{ab-initio} Hamiltonian 
\begin{equation}
\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 -\frac{1}{2\mu}\sum_{I} \nabla_I^2 + \frac{1}{2}\sum_{ij} \frac{1}{|r_i - r_j|} + \frac{1}{2}\sum_{IJ} \frac{Z_I Z_J}{|R_I - R_J|} - \sum_{iJ}\frac{Z_J}{|r_i - R_J|},\ \mu = (M/m)
\label{Hab}
\end{equation}
where $i$ index electrons, $I$ ions, and energy in Hartree (Ha).
The procedure makes no assumptions about the form of microscopic electronic interactions in the system and reports a clear, statistical measure of model quality.
By developing many-body models for SLG and BLG, I hope to further the understanding of interactions in graphene like systems and the maturity of DMD for its application to other strongly correlated systems like TBLG.

@@

A new constrained variational Monte Carlo (cVMC) method will be used to sample the low-energy wave functions necessary for DMD.
The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
\begin{equation}
\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \vec{\lambda} \cdot \langle \Psi(\vec{p}) | (\vec{d} - \vec{d}_0)^2 | \Psi(\vec{p}) \rangle 
\end{equation}
with respect to the parameters $\vec{p}$.
The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}_0$ is the target values of those candidate descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}_0$.
Essentially this method allows us, for a given parameterization of $|\Psi(\vec{p})\rangle$, to generate low-energy wave functions which sit near a target point $\vec{d}_0$ in candidate descriptor space.

@@

Apart from their individual value, the four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
The first question would be whether the parameter values between a finite size system, like benzene, match those seen in the bulk SLG system.
The second would be whether the intralayer parameter values change in each sheet of SLG when they are brought together in the AA and AB configuration.
The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large \cite{Bistritzer2011}.

>>>>

\begin{document}
\maketitle
\begin{abstract}
A core problem of condensed matter physics is understanding how to develop low-energy effective models for realistic systems.
I believe that a density matrix downfolding (DMD) method \cite{Zheng2017} can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian.
I present as a preliminary result an effective Hamiltonian developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
I propose using DMD to develop accurate model Hamiltonians for single and bilayer graphene.
My work will serve to mature the DMD method and take steps towards a model for the strongly correlated twisted bilayer graphene system.
\end{abstract}
Committee: Lucas Wagner, Brian DeMarco, Andre Schleife, Karin Dahmen
\pagebreak

@@

A relatively new technique is density matrix downfolding (DMD), which provides a framework for developing many-body models directly from calculations on the \textit{ab-initio} Hamiltonian
\begin{equation}
\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 -\frac{1}{2\mu}\sum_{I} \nabla_I^2 + \frac{1}{2}\sum_{ij} \frac{1}{|r_i - r_j|} + \frac{1}{2}\sum_{IJ} \frac{Z_I Z_J}{|R_I - R_J|} - \sum_{iJ}\frac{Z_J}{|r_i - R_J|},\ \mu = (M/m)
\label{Hab}
\end{equation}
where $i$ index electrons, $I$ ions, and energy in Hartree (Ha).
The procedure makes no assumptions about the form of microscopic electronic interactions in the system and reports a clear, statistical measure of model quality.
I propose using a quantum Monte Carlo (QMC) DMD procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
Recent observations of a fractional quantum Hall effect \cite {Bolotin2009, Du2009} and plasmarons \cite{Bostwick999} in SLG as well as unconventional superconductivity \cite{Cao2018_2} and many-body insulating states \cite{Cao2018} in TBLG have spurred an interest in developing many-body models for graphene.
By developing many-body models for SLG and BLG, I hope to further the understanding of interactions in graphene like systems and the maturity of DMD for its application to other strongly correlated systems like TBLG.

@@

A new constrained variational Monte Carlo (cVMC) method will be used to sample the low-energy wave functions necessary for DMD.
The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
\begin{equation}
\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2
\end{equation}
with respect to the parameters $\vec{p}$.
The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^0$ is the target values of those candidate descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}_0$.
Essentially this method allows us, for a given parameterization of $|\Psi(\vec{p})\rangle$, to generate low-energy wave functions which sit near a target point $\vec{d}^0$ in candidate descriptor space.

@@

Apart from their individual value, the four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
The first question would be whether the parameter values between a finite size system, like benzene, match those seen in the bulk SLG system.
The second would be whether the intralayer parameter values change in each sheet of SLG when they are brought together in the AA and AB configuration.
The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large \cite{Bistritzer2	011}.
